---
title: "grocery_eda"
author: "Meghna Diwan"
date: "5/21/2020"
output: html_document
---

```{r}
library(data.table)
library(tidyverse)
library(lubridate)
library(xts)
library(leaflet)
library(ggfortify)
library(tidyquant)

library(forecast)
library(tseries)
library(TSA)
rm(list = ls())
gc()
```


# Import datasets
```{r}
datapath = "/Volumes/SSD/TS-Project/data"
dtrain <- fread(file=paste(datapath,"full/train.csv",sep="/"))
stores <- fread(file=paste(datapath,"full/stores.csv",sep="/"))
#oil <- fread(file=paste(datapath,"oil.csv",sep="/"))
items <- fread(file=paste(datapath,"full/items.csv",sep="/"))
transactions <- fread('/Volumes/SSD/TS-Project/favorita-grocery-sales-forecasting/transactions.csv')
#holiday = fread('/Volumes/SSD/TS-Project/favorita-grocery-sales-forecasting/holidays_events.csv')
#test = fread('/Volumes/SSD/TS-Project/favorita-grocery-sales-forecasting/test.csv')

library(readxl)
item_map = read_excel(path = paste(datapath,"full/store_item_map.xlsx",sep="/"), sheet = "item_map")
store_map = read_excel(path = paste(datapath,"full/store_item_map.xlsx",sep="/"), sheet = "store_map")
cluster_map = read_excel(path = paste(datapath,"full/store_item_map.xlsx",sep="/"), sheet = "cluster_map")
```

# Add Item and Store maps to iten and store df
```{r}
items = left_join(items, item_map, by= 'family')
stores = left_join(stores, store_map, by = 'store_nbr')
stores = left_join(stores, cluster_map, by = 'cluster')
```

# Check for missing values
```{r}
data.frame(dtrain = colSums(is.na(dtrain)))
data.frame(stores = colSums(is.na(stores)))
data.frame(oil = colSums(is.na(oil))) # need to remove nas
data.frame(items = colSums(is.na(items)))
data.frame(transactions = colSums(is.na(transactions)))
data.frame(holiday = colSums(is.na(holiday)))
data.frame(test = colSums(is.na(test)))
```

# Data
```{r}
glimpse(dtrain)
unique(items$family)
unique(dtrain$store_nbr)
unique(stores$cluster) # groups of similar stores
```

# Top 10 stores by item sold
```{r}
Top10Stores = dtrain %>%
  group_by(store_nbr) %>%
  summarise(Count = sum(unit_sales)) %>%
  arrange(desc(Count)) %>%
  head(10) 

Top10StoresDescription = inner_join(Top10Stores,stores)
Top10StoresDescription
```

# Top Selling items in Store 44
```{r}
Top10Items = dtrain %>%
                filter(store_nbr == 44) %>%
                filter(!is.na(item_nbr)) %>%
                group_by(item_nbr) %>%
                summarise(Count = n()) %>%
                ungroup() %>%
                mutate(item_nbr = reorder(item_nbr,Count)) %>%
                arrange(desc(Count)) %>%
                head(10)
Top10Items$item_nbr = as.integer(as.character(Top10Items$item_nbr))
Top10ItemsDescription = inner_join(Top10Items,items)
Top10ItemsDescription
```

# Check if all dates are available between 2013 and 2017
```{r}
dates_all = seq(as.Date("2013-01-01"), as.Date("2017-08-15"), by="days")
dates = as.Date(unique(dtrain$date))
as.Date(setdiff(dates_all, dates)) # dates missing are Christmas day in all 4 years - stores closed on that date

# Oil Dates - not continuous
head(oil$date)
tail(oil$date)
oil$date = as.Date(oil$date)
as.Date(setdiff(dates_all, oil$date))
```

# Visualize Dates
```{r}
train_dt <- dtrain %>%
  distinct(date) %>%
  mutate(dset = "train") %>%
  mutate(year = year(date))

train_dt$date = ymd(as.Date(train_dt$date, "%Y-%m-%d"))
year(train_dt$date) = 2017

train_dt %>%
  filter(!is.na(date)) %>%
  ggplot(aes(date, year)) +
  geom_point(shape = "|", size = 10, colour = "#2a3990") +
  scale_x_date(date_labels = "%B", date_breaks = "1 month") +
  scale_y_reverse() +
  theme(axis.text.x  = element_text(angle=45, hjust=1, vjust=0.9))
```


# Perishable
```{r}
items %>%
  ggplot(aes(family, fill = perishable)) +
  geom_bar(position = "fill") +
  scale_fill_continuous(type = "viridis") +
  coord_flip() 


items %>% 
  ggplot(aes(family, perishable)) +  
  geom_bar(stat = "identity", na.rm = T, colour = "#2a3990") + 
  coord_flip()
```


# Promotions
```{r}
item_promo = dtrain %>% 
                select(item_nbr, promotion) %>% 
                group_by(item_nbr) %>% 
                summarize(n_promo = sum(promotion, na.rm = T))

items = left_join(items, item_promo)
items$n_promo[is.na(items$n_promo)] <- 0


items %>% 
  ggplot(aes(family, n_promo)) +  
  geom_bar(stat = "identity", na.rm = T, colour = "#2a3990") + 
  coord_flip() 
```


# Collapse Data
```{r}
# Shopping behavior patterns of Ecudor
dtrain$promotion = ifelse(dtrain$onpromotion == TRUE, 1, 0)
dtrain$promotion[is.na(dtrain$promotion)] = 0

dtrain_item <- left_join(dtrain, items)

dtrain_item_lvl = dtrain_item %>%
                    select(date, item_nbr, family, unit_sales, promotion, perishable) %>%
                    group_by(date, family) %>%
                    summarize(tot_sales = sum(unit_sales), 
                              promotion = max(promotion, na.rm = T),
                              perishable = sum(perishable))

TopItems = dtrain_item_lvl %>%
                group_by(family) %>%
                summarise(sales = sum(tot_sales)) %>%
                ungroup() %>%
                mutate(family = reorder(family,sales)) %>%
                arrange(desc(sales)) %>%
                head(10)
TopItems
```

# Create Time Series
```{r}
grocery = subset(dtrain_item_lvl, select = tot_sales,
                 family == 'GROCERY I')
grocery_ts = ts(grocery, frequency=365.25, 
     start = c(year("2013-01-01"), 1))
plot(grocery_ts, col = "blue")
acf(grocery_ts, 100)
pacf(grocery_ts, 100)
eacf(grocery_ts)
```

# Add Christmas day back to dataset - maybe?
```{r}
fam = unique(items$family)
dat = c("2013-12-25", "2014-12-25", "2015-12-25", "2016-12-25")
xmas = c()
i = 1
for (f in fam) {
  for (d in dat) {
    xmas = rbind(xmas, c('date' = d, 'family' = f, 'tot_sales' = 0, 'promotion' = 0, 'perishable' = 0))
  }
}
```


# Prep for Heirachical TS
```{r}
train <- left_join(dtrain, items)
train_fam = train %>%
              select(date, store_nbr, item_nbr, family, family_nbr, unit_sales) %>%
              group_by(date, store_nbr, family, family_nbr) %>%
              summarize(tot_sales = sum(unit_sales))
train_fam = left_join(train_fam, stores)
write.csv(train_fam, paste(datapath,"subset/train_store_fam.csv",sep="/"))
```

# Create HTS TS
https://pkg.earo.me/hts/
```{r}
library(hts)
library(fpp2)
library(tsbox)


train = fread(file=paste(datapath,"subset/train_store_fam.csv",sep="/"))
subset = train %>% 
  select(date, store_char, cluster_nbr, family_nbr, tot_sales) %>%
  mutate(date = as.Date(date, format = "%Y-%m-%d")) %>%
  #filter(date >= as.Date('2013-01-01') & date < as.Date('2014-01-01')) %>%
  filter(family_nbr %in% c("fam13", "fam04", "fam08", "fam31", "fam09", "fam26")) %>% 
  # Grocery I, Beverages, Cleaning, Produce, Dairy & Personal Care
  filter(cluster_nbr %in% c("c01", "c02", "c03", "c04"))


# family and cluster
subset$id = paste0(subset$cluster_nbr,subset$family_nbr)
subset_wide = spread(subset, id, tot_sales)
subset_wide[is.na(subset_wide)] = 0
subset_wide = subset_wide %>%
                  select(-store_char, -cluster_nbr, -family_nbr) %>%
                  group_by(date) %>%
                  summarise_all(funs(sum))
subset.ts = as.xts(subset_wide[2:ncol(subset_wide)],order.by=as.Date(subset_wide$date, "%Y-%m-%d")) 

# Find nodes for each level
n_nodes = subset %>%
            group_by(cluster_nbr) %>%
            summarise(n_fam = COUNT_UNIQUE(family_nbr))

# Create HTS
subset_hts = hts(subset.ts, characters = c(3, 5))
summary(subset_hts)
smatrix(subset_hts)
fcst<-forecast(subset_hts, method="mo", fmethod="arima", level = 1, h=10)

# Graph by stores
subset_hts %>% aggts(levels=0:1) %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Sales") + ggtitle("Grocery Stores")

# Graph by Store and Item Family
cols <- sample(scales::hue_pal(h=c(15,375),
  c=100,l=65,h.start=0,direction = 1)(NCOL(subset.ts)))
as_tibble(subset.ts) %>%
  gather(Items) %>%
  mutate(Date = rep(time(subset.ts), NCOL(subset.ts)),
         State = str_sub(Items,1,3)) %>%
  ggplot(aes(x=Date, y=value, group=Items, colour=Items)) +
    geom_line() +
    facet_grid(State~., scales="free_y") +
    xlab("Year") + ylab("Sales") +
    ggtitle("Grocery Sales by Item Family") +
    scale_colour_manual(values = cols)
```

## HTS - Version 2
```{r}
top_item = train %>%
      filter(city %in% c("Quito", "Guayaquil")) %>%
      group_by(city, type, family) %>%
      summarize(tot_sales = sum(tot_sales)) %>%
      arrange(city, type, desc(tot_sales)) %>%
      top_n(5, tot_sales)
  
train_hts = inner_join(top_item, train, by= c("city", "type", "family"))
train_hts = train_hts %>%
                select(date, city, type, family, family_nbr, tot_sales.y) %>%
                group_by(date, city, type, family, family_nbr) %>%
                summarise(tot_sales = sum(tot_sales.y))
train_hts$city_label = substr(train_hts$city, 1, 2)
train_hts$label = paste0(train_hts$city_label,train_hts$type,train_hts$family_nbr)
subset_wide = spread(subset, id, tot_sales)
subset_wide[is.na(subset_wide)] = 0
subset_wide = subset_wide %>%
                  select(-store_char, -cluster_nbr, -family_nbr) %>%
                  group_by(date) %>%
                  summarise_all(funs(sum))
subset.ts = as.xts(subset_wide[2:ncol(subset_wide)],order.by=as.Date(subset_wide$date, "%Y-%m-%d")) 
```


## Auto.Arima
```{r}
grocery = subset(train, select = tot_sales,
                 family == 'GROCERY I' & store_nbr ==12)
grocery_ts = ts(grocery, frequency=365.25, 
     start = c(year("2013-01-01"), 1))
plot(grocery_ts, col = "blue")
acf(grocery_ts, 100)
pacf(grocery_ts, 100)
eacf(grocery_ts)

arima_grocery = auto.arima(grocery_ts, seasonal = T)
pred = forecast(arima_grocery, h = 60)
plot(pred)
```

## Most Transactions
```{r}
trans_summ = transactions %>%
                group_by(store_nbr) %>%
                summarize(tot_transactions = sum(transactions))

store_info = left_join(stores, trans_summ)
```

```{r}
tot_sales = train %>%
            select(date, tot_sales) %>%
            group_by(date) %>%
            summarise(sales = sum(tot_sales)) %>%
            filter(date > as.Date('2015-01-01'))
sales.xts = as.xts(tot_sales$sales,order.by=as.Date(tot_sales$date, "%Y-%m-%d"))
dev.new(width=30, height=9, unit="in")
plot(sales.xts)
```


# Extra Code
```{r}
library(MTS)
library()

# store1
store_1 = subset(train_store_item, store_nbr == 1)
store_1 = spread(store_1, family, tot_sales)
store_1[is.na(store_1)] = 0
store_1.ts = as.xts(store_1[3:34],order.by=as.Date(store_1$date, "%Y-%m-%d"))

# store2
store_2 = subset(train_store_item, store_nbr == 2)
store_2 = spread(store_2, family, tot_sales)
store_2[is.na(store_2)] = 0
store_2.ts = as.xts(store_2[3:34],order.by=as.Date(store_2$date, "%Y-%m-%d"))
```
