---
title: "Preprocess Holidays"
author: "Aarti M Rao"
date: "5/24/2020"
output: html_document
---
```{r setup, include=FALSE}
library(dplyr)
```

# Import data
```{r}
holidays <- read.csv('holidays_events.csv')
#convert the date column to date format
holidays$date <- as.Date(holidays$date)
```

# Remove dates with type="Work Day"
```{r}
holidays_nowd <- holidays %>%
  filter(type != "Work Day")
```

# Remove dates with transferred = "True"
```{r}
holidays_notransfer <- holidays_nowd %>%
  filter(transferred == "False")
```

# Remove unnecessary columns (description, transferred, locale name)
```{r}
# using unique function to remove duplicate records
holidays_nodup <- unique(holidays_notransfer[ ,c(1,3)])
```

# Transform locale column to be a binary flag to indicate whether the holiday was National or not
```{r}
# replacing "Local" with "Regional"
holidays_loc <- holidays_nodup %>%
  mutate(locale=replace(locale, locale=="Local", "Regional")) %>%
  as.data.frame()

# replacing values with 1s(National) and 0s (Regional)
holidays_loc$locale=ifelse(holidays_loc$locale=="National",1,0)

#  Grouping by the date to remove duplicates - if a date captured a holiday once as a regional and once as a National Holiday, its aggregated as a National Holiday
holidays_unique = holidays_loc %>%
  group_by(date) %>%
  summarize(national_flag = max(locale))

summary(holidays_unique)
```

# Trimming the holiday dates range to within train data date range (as dates start from 02 Mar 2012 and go up tp 26 dec 2017)
```{r}
holidays_unique <- subset(holidays_unique, date>= "2013-01-01" & date < "2017-09-01")
summary(holidays_unique)
```

# Creating continuous sequence of dates
```{r}
dates <- data.frame(seq(as.Date("2013/1/1"), as.Date("2017/8/31"), "day"))
names(dates) <- "date"
```

# Join holidays_unique to dates sequence
```{r}
holidays_all <- left_join(dates, holidays_unique)
```

# Creating a holiday flag
```{r}
holidays_all$holiday_flag=ifelse(is.na(holidays_all$national_flag),0,1)
```

# Replacing NAs in Nationa flag with 0s
```{r}
holidays_all$national_flag[is.na(holidays_all$national_flag)] <- 0
summary(holidays_all)
```

# Saving the dataset
```{r}
write.csv(holidays_all,'holidays_new.csv')
```