---
title: "Preprocess Oil Prices"
author: "Aarti Rao, Grace Chai, Jenny Huang, Meghna Diwan"
date: "5/24/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(zoo)
```

# Import data
```{r}
datapath = "/Volumes/SSD/TS-Project/data"
oil = read.csv(file = paste(datapath, 'full/oil.csv', sep = "/"))
#convert the date column to date format
oil$date <- as.Date(oil$date)
```

# Creating continuous sequence of dates
```{r}
dates <- data.frame(seq(as.Date("2013/1/1"), as.Date("2017/8/31"), "day"))
names(dates) <- "date"
```

# Join oil prices to dates sequence to identify the dates with missing prices
```{r}
oil_all <- left_join(dates, oil)
```

# Fill NAs with prices from previous fields
## Logic - it appears that weekends and holidays have oil prices missing. Filling them with the closest available price as they wouldn't change over the weekend
```{r}
# using the na.locf(last observation forward) function from the zoo package
oil_all$price <- na.locf(oil_all$dcoilwtico, na.rm=FALSE)
```

# Carrying forward only the required columns
```{r}
oil_new <- oil_all[c("date", "price")]
#filling the missing value for 01-Jan-2013 with value from 02-Jan-2013
oil_new[1,2] <- oil_new[2,2]
```

# Saving the dataset
```{r eval=FALSE}
write.csv(oil_new,'oil_new.csv')
```

# Math to verify number of dates with missing prices
## More or less equal to (no. of weeks * 2) (number of weekend days) in ~4.6 years data

```{r}
#1704-1218

#(4*365)+243
#(365/12)*8
#(((244/365)+4)*52)*2
```

