---
title: "Hierarchical Time Series Models"
author: "Aarti Rao, Grace Chai, Jenny Huang, Meghna Diwan"
date: "5/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, echo=TRUE, warning=FALSE, message=FALSE, fig.keep = 'all')
```

```{r message=FALSE, warning=FALSE}
library(data.table)
library(tidyverse)
library(lubridate)
library(xts)
library(leaflet)
library(ggfortify)
library(tidyquant)
library(forecast)
library(tseries)
library(TSA)
library(hts)
library(fpp2)
library(tsbox)
rm(list = ls())
gc()
```

# Import Data
```{r}
datapath = "/Volumes/SSD/TS-Project/data"
train = fread(file=paste(datapath,"subset/train_city_type_item.csv",sep="/"))
train = fread('/Users/meghnadiwan/Desktop/train_city_type_item.csv')
```

# Transform to Wide form
```{r}
# Long to Wide
train_wide = spread(train, hts_label, tot_sales)
train_wide[is.na(train_wide)] = 0
train_wide = train_wide %>%
                  select(-city, -type, -family, -price, -n_promotion, -perishable, -national_flag,
                         -holiday_flag, -pay.day_flag) %>%
                  group_by(date) %>%
                  summarise_all(funs(sum))
# Covert to TS object
train.ts = ts(train_wide[2:ncol(train_wide)], frequency=365.25, 
     start = c(year("2013-01-01"), 1))
```

# Run HTS
```{r}
train.hts = hts(train.ts, characters = c(2, 1, 5))
summary(train.hts)
smatrix(train.hts)
```

# Plot HTS
```{r}
# Graph by City
train.hts %>% aggts(levels=0:1) %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Total Sales") + ggtitle("Grocery Sales by City")
```

```{r}
# Graph by Type - City Guayaquil
aggts(train.hts, levels=1:2)[,c(1,3:7)] %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Total Sales") + ggtitle("Grocery Sales by City & Type - Guayaquil")
```

```{r}
# Graph by Type - City Quito
aggts(train.hts, levels=1:2)[,c(2,8:11)] %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Total Sales") + ggtitle("Grocery Sales by City & Type - Quito")
```

```{r}
# Graph by Item Family - Type A, City Guayaquil
aggts(train.hts, levels=2:3)[,c(1,10:14)] %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Total Sales") + ggtitle("Grocery Sales by City, Type & Item Family - Guayaquil Type A")
```

```{r}
# Graph by Item Family - Type B, City Guayaquil
aggts(train.hts, levels=2:3)[,c(2,15:19)] %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Total Sales") + ggtitle("Grocery Sales by City, Type & Item Family - Guayaquil Type B")
```

```{r}
# Graph by Item Family - Type C, City Guayaquil
aggts(train.hts, levels=2:3)[,c(3,20:24)] %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Total Sales") + ggtitle("Grocery Sales by City, Type & Item Family - Guayaquil Type C")
```

```{r}
# Graph by Item Family - Type D, City Guayaquil
aggts(train.hts, levels=2:3)[,c(4,25:29)] %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Total Sales") + ggtitle("Grocery Sales by City, Type & Item Family - Guayaquil Type D")
```

```{r}
# Graph by Item Family - Type E, City Guayaquil
aggts(train.hts, levels=2:3)[,c(5,30:34)] %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Total Sales") + ggtitle("Grocery Sales by City, Type & Item Family - Guayaquil Type E")
```

```{r}
# Graph by Item Family - Type A, City Quito
aggts(train.hts, levels=2:3)[,c(6,35:39)] %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Total Sales") + ggtitle("Grocery Sales by City, Type & Item Family - Quito Type A")
```

```{r}
# Graph by Item Family - Type B, City Quito
aggts(train.hts, levels=2:3)[,c(7,40:44)] %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Total Sales") + ggtitle("Grocery Sales by City, Type & Item Family - Quito Type B")
```

```{r}
# Graph by Item Family - Type C, City Quito
aggts(train.hts, levels=2:3)[,c(8,45:49)] %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Total Sales") + ggtitle("Grocery Sales by City, Type & Item Family - Quito Type C")
```

```{r}
# Graph by Item Family - Type D, City Quito
aggts(train.hts, levels=2:3)[,c(9,50:54)] %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Total Sales") + ggtitle("Grocery Sales by City, Type & Item Family - Quito Type D")

```

# Train - Val Split
```{r}
data.hts = window(train.hts, start = 2013, end = 2017.541) # 1659 dates
summary(data.hts)
val.hts = window(train.hts, start = 2017.542) # 20 dates
summary(val.hts)
```

# SMAPE Function
```{r}
smape = function(y_pred, y_actual) {
  error = mean(abs(y_pred - y_actual)/(abs(y_actual) + abs(y_pred)))
  return(error)
}
```

# Base Model
```{r}
rw.top.down = forecast(
  data.hts, h = 20, method = "tdfp", fmethod = "rw",
  keep.fitted = TRUE, keep.resid = TRUE
)

rw.bottom.up = forecast(
  data.hts, h = 20, method = "bu", fmethod = "rw",
  keep.fitted = TRUE, keep.resid = TRUE
)

rw.middle.out = forecast(
  data.hts, h = 20, method = "mo", fmethod = "rw", level = 2,
  keep.fitted = TRUE, keep.resid = TRUE
)

rw.comb = forecast(
  data.hts, h = 20, method = "comb", fmethod = "rw",
  keep.fitted = TRUE, keep.resid = TRUE
)

base_results = rbind(smape(rw.top.down$bts, val.hts$bts),
                smape(rw.bottom.up$bts, val.hts$bts),
                smape(rw.middle.out$bts, val.hts$bts),
                smape(rw.comb$bts, val.hts$bts))
rownames(base_results) = c("top.down",  "bottom.up",  "middle.out", "combination")
base_results
```
    
Irrespective of method, hierarchical time series with random walk gives the same SMAPE.

# Base model plots (Top Down)
```{r}
rw.top.down$bts[,1:5] %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Total Sales") + ggtitle("Forecasted Grocery Sales by City, Type & Item Family - Guayaquil Type A")+
  ylim(800,10000)

val.hts$bts[,1:5] %>%
  autoplot(facet=TRUE, color='red')+
  xlab("Year") + ylab("Total Sales") + ggtitle("Actual Grocery Sales by City, Type & Item Family - Guayaquil Type A")+
  ylim(800,10000)
```

# Base model plots (Top Down)
```{r}
rw.top.down$bts[,26:30] %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Total Sales") + ggtitle("Forecasted Grocery Sales by City, Type & Item Family - Quinto Type A")+
  ylim(8000,87000)

val.hts$bts[,26:30] %>%
  autoplot(facet=TRUE, color='red')+
  xlab("Year") + ylab("Total Sales") + ggtitle("Actual Grocery Sales by City, Type & Item Family - Quinto Type A")+
  ylim(8000,87000)

```

# Forecast using different methods
```{r}
top.down = forecast(
  data.hts, h = 20, method = "tdfp", fmethod = "arima",
  keep.fitted = TRUE, keep.resid = TRUE
)

bottom.up = forecast(
  data.hts, h = 20, method = "bu", fmethod = "arima",
  keep.fitted = TRUE, keep.resid = TRUE
)

middle.out = forecast(
  data.hts, h = 20, method = "mo", fmethod = "arima", level = 2,
  keep.fitted = TRUE, keep.resid = TRUE
)

comb = forecast(
  data.hts, h = 20, method = "comb", fmethod = "arima",
  keep.fitted = TRUE, keep.resid = TRUE
)

comb_wls = forecast(
  data.hts, h = 20, method = "comb", fmethod = "arima", weights = "wls",
  keep.fitted = TRUE, keep.resid = TRUE
)
 
comb_wls_alg = forecast(
  data.hts, h = 20, method = "comb", fmethod = "arima", weights = "wls", algorithms = "lu",
  keep.fitted = TRUE, keep.resid = TRUE
)

comb_nseries = forecast(
  data.hts, h = 20, method = "comb", fmethod = "arima", weights = "nseries",
  keep.fitted = TRUE, keep.resid = TRUE
)

```

# Compare SMAPE Errors
```{r}
results = rbind(smape(top.down$bts, val.hts$bts),
                smape(bottom.up$bts, val.hts$bts),
                smape(middle.out$bts, val.hts$bts),
                smape(comb$bts, val.hts$bts),
                smape(comb_wls$bts, val.hts$bts),
                smape(comb_wls_alg$bts, val.hts$bts), 
                smape(comb_nseries$bts, val.hts$bts))
rownames(results) = c("top.down",  "bottom.up",  "middle.out", 
                      "combination", "comb_wls", "comb_wls_alg", "comb_nseries")
results
```

# Forecast Plots - Arima 
```{r}
comb$bts[,1:5] %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Total Sales") + ggtitle("Forecasted Grocery Sales by City, Type & Item Family - Guayaquil Type A")+
  ylim(800,10000)

val.hts$bts[,1:5] %>%
  autoplot(facet=TRUE, color='red')+
  xlab("Year") + ylab("Total Sales") + ggtitle("Actual Grocery Sales by City, Type & Item Family - Guayaquil Type A")+
  ylim(800,10000)

```


# Forecast Plots - Arima 
```{r}
comb$bts[,26:30] %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Total Sales") + ggtitle("Forecasted Grocery Sales by City, Type & Item Family - Quinto Type A")+
  ylim(8000,90000)

val.hts$bts[,26:30] %>%
  autoplot(facet=TRUE, color='red')+
  xlab("Year") + ylab("Total Sales") + ggtitle("Actual Grocery Sales by City, Type & Item Family - Quinto Type A")+
  ylim(8000,90000)

```

# Forecast - Arima with XREG
```{r}
oil_price <- unique(train[,c("date","price")])
train_oil_price <- oil_price$price[1:1659]
test_oil_price <- oil_price$price[1660:1679]
```

```{r}
comb_oil_xreg = forecast(
  data.hts, h = 20, method = "comb", fmethod = "arima",
  keep.fitted = TRUE, keep.resid = TRUE, xreg = train_oil_price, newxreg = test_oil_price
)
```

```{r}
smape(comb_oil_xreg$bts, val.hts$bts)
```

```{r}
x_var <- unique(train[,c(2,7,10:12)])
train_x_var <- as.matrix(x_var[1:1659,2:5])
test_x_var <- as.matrix(x_var[1660:1679,2:5])
```

```{r}
comb_all_xreg = forecast(
  data.hts, h = 20, method = "comb", fmethod = "arima",
  keep.fitted = TRUE, keep.resid = TRUE, xreg = train_x_var, newxreg = test_x_var
)
```

```{r}
smape(comb_all_xreg$bts, val.hts$bts)
```

# Forecast Plots - Arima +Xreg
```{r}
comb_all_xreg$bts[,1:5] %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Total Sales") + ggtitle("Forecasted Grocery Sales by City, Type & Item Family - Guayaquil Type A")+
  ylim(800,10000)

val.hts$bts[,1:5] %>%
  autoplot(facet=TRUE, color='red')+
  xlab("Year") + ylab("Total Sales") + ggtitle("Actual Grocery Sales by City, Type & Item Family - Guayaquil Type A")+
  ylim(800,10000)

```

# Forecast Plots - Arima Xreg 
```{r}
comb_all_xreg$bts[,26:30] %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Total Sales") + ggtitle("Forecasted Grocery Sales by City, Type & Item Family - Quinto Type A")+
  ylim(8000,90000)

val.hts$bts[,26:30] %>%
  autoplot(facet=TRUE, color='red')+
  xlab("Year") + ylab("Total Sales") + ggtitle("Actual Grocery Sales by City, Type & Item Family - Quinto Type A")+
  ylim(8000,90000)

```

# 
```{r}
# total n_promotions in unique day
total_n_promotion = train %>% dplyr::group_by(date) %>% dplyr::summarise(total_n_promotion =sum(n_promotion))
## total number of perishable
total_perishable = train %>% dplyr::group_by(date) %>% dplyr::summarise(total_perishable = sum(perishable))

x_var$total_n_promotion <- total_n_promotion$total_n_promotion
x_var$total_perishable <- total_perishable$total_perishable


train_x_var_all <- as.matrix(x_var[1:1659,2:7])
test_x_var_all <- as.matrix(x_var[1660:1679,2:7])
```

```{r}
comb_all_xreg_n = forecast(
  data.hts, h = 20, method = "comb", fmethod = "arima",
  keep.fitted = TRUE, keep.resid = TRUE, xreg = train_x_var_all, newxreg = test_x_var_all
)
```

```{r}
smape(comb_all_xreg_n$bts, val.hts$bts)
```


```{r}
comb_all_xreg_n$bts[,1:5] %>%
  autoplot(facet=TRUE) +
  xlab("Year") + ylab("Total Sales") + ggtitle("Forecasted Grocery Sales by City, Type & Item Family - Guayaquil Type A")+
  ylim(800,10000)

val.hts$bts[,1:5] %>%
  autoplot(facet=TRUE, color='red')+
  xlab("Year") + ylab("Total Sales") + ggtitle("Actual Grocery Sales by City, Type & Item Family - Guayaquil Type A")+
  ylim(800,10000)

```


```{r}
smape.values <- data.frame()
for (i in 1:45){
  smape.values<-rbind(smape.values, smape(val.hts$bts[,i],comb_all_xreg_n$bts[,i]))
}

smape.values

```

```{r}
smape.final <- colSums(smape.values)/45; smape.final
```

# Cross-Validation Function
```{r}
hts.CV = function(y, window, h) {
  cv.error = matrix(ncol = 1)
  
  for (i in seq(1, nrow(y) - (window + h) + 1)) {
    train_start = i
    train_end = i + window - 1
    test_start = i + window
    test_end = i + window + h - 1
   
    train_y = subset(y, start = train_start, end = train_end)
    test_y = subset(y, start = test_start, end = test_end)

    train_hts = hts(train_y, characters = c(2, 1, 5))
    test_hts = hts(test_y, characters = c(2, 1, 5))
    
    train_x_var_all_cv <- as.matrix(x_var[train_start:train_end,2:7])
    test_x_var_all_cv <- as.matrix(x_var[test_start:test_end,2:7])
    
    fc = forecast(train_hts, h = h, method = "comb", fmethod = "arima",
    keep.fitted = TRUE, keep.resid = TRUE, xreg = train_x_var_all_cv, newxreg = test_x_var_all_cv)
    print(paste("Done", as.character(i)))
    acc = smape(fc$bts, test_hts$bts)
    cv.error = rbind(cv.error, acc)
  }
  
  return(cv.error)
}
```

# CV Using 10 Folds
```{r}
start.time = Sys.time()
cv.acc = hts.CV(y = train.ts, window = 1640, h = 30) # takes 11 mins to run one fold
end.time = Sys.time()
(time.taken = end.time - start.time)
```
    
Due to limited computational power, only 10-fold CV was implemented.

# CV Errors
```{r}
# Mean SMAPE
mean(cv.acc, na.rm = T)

# Standard Deviation of SMAPE
sd(cv.acc, na.rm = T)
```
